---
title: "meteoland - Landscape meteorology tools"
output: github_document
---

```{r setup, echo = FALSE}

knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "README-")
```

  <!-- badges: start -->
  [![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/meteoland)](https://cran.r-project.org/package=meteoland)
  [![](https://cranlogs.r-pkg.org/badges/meteoland)](https://cran.rstudio.com/web/packages/meteoland/index.html)
  [![R-CMD-check](https://github.com/emf-creaf/meteoland/workflows/R-CMD-check/badge.svg)](https://github.com/emf-creaf/meteoland/actions)
  <!-- badges: end -->

## Introduction

With the aim to assist research of climatic impacts on forests, the R package `meteoland` provides utilities to estimate daily weather variables at any position over complex terrains:

+ Spatial interpolation of daily weather records from meteorological stations.
+ Statistical correction of meteorological data series (e.g. from climate models).
+ Multisite and multivariate stochastic weather generation.

A more detailed introduction to the package functionality can be found in De Cáceres et al. (2018).

## Package installation and documentation

Package `meteoland` can be found at [CRAN](https://cran.r-project.org/), but the version in this repository may not be the most recent one. Latest stable versions can be downloaded and installed from GitHub as follows (package `remotes` should be installed first):

```{r installation, eval = FALSE}
remotes::install_github("emf-creaf/meteoland")
```

Alternatively, users can have help to run package functions directly as package vignettes, by forcing their inclusion in installation: 
```{r installation vignettes, eval = FALSE}
remotes::install_github("emf-creaf/meteoland", 
                        build_opts = c("--no-resave-data", "--no-manual"),
                        build_vignettes = TRUE)
```

Detailed documentation on `meteoland` calculation routines can be found at (https://emf-creaf.github.io/meteolandbook/index.html).

## Companion packages

During the development of `meteoland` some functions to download weather station data from several Spanish networks were originally developed. After `meteoland` version 1.0.1, the user is recommended to use package [**meteospain**](https://emf-creaf.github.io/meteospain/), which can also be found at [CRAN](https://cran.rstudio.com/web/packages/meteospain/index.html). Functions to download weather station data are still available in `meteoland` but they have been deprecated and make internal calls to functions in package [**meteospain**](https://emf-creaf.github.io/meteospain/).

## References

+ De Caceres M, Martin-StPaul N, Turco M, Cabon A, Granda V (2018) Estimating daily meteorological data and downscaling climate models over landscapes. Environmental Modelling and Software 108: 186-196. (doi:10.1016/j.envsoft.2018.08.003).

