---
title: "meteoland - Landscape meteorology tools"
output: github_document
---

```{r setup, echo = FALSE}

knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "README-")
```

## Introduction

With the aim to assist research of climatic impacts on forests, the R package `meteoland` provides utilities to estimate daily weather variables at any position over complex terrains:

+ Spatial interpolation of daily weather records from meteorological stations.
+ Statistical correction of meteorological data series (e.g. from climate models).

A more detailed introduction to the package functionality can be found in De Cáceres et al. (2018).

## Package installation and documentation

Package `meteoland` can be found at [CRAN](https://cran.r-project.org/), but the version in this repository may not be the most recent one. Latest stable versions can be downloaded and installed from GitHub as follows (package `devtools` should be installed first):

```{r installation, eval = FALSE}
devtools::install_github("vegmod/meteoland")
```

Alternatively, users can have help to run package functions directly as package vignettes, by forcing their inclusion in installation: 
```{r installation vignettes, eval = FALSE}
devtools::install_github("vegmod/meteoland", 
                         build_opts = c("--no-resave-data", "--no-manual"))
```

Detailed documentation on `meteoland` can be found at (https://vegmod.ctfc.cat/software/meteoland).

## References

+ De Caceres M, Martin-StPaul N, Turco M, Cabon A, Granda V (2018) Estimating daily meteorological data and downscaling climate models over landscapes. Environmental Modelling and Software 108: 186-196. (doi:10.1016/j.envsoft.2018.08.003).

