---
title: "Function equivalences between older (< 2.0.0) and newer (>= 2.0.0) meteoland versions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Function equivalences between older (< 2.0.0) and newer (>= 2.0.0) meteoland versions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(meteoland)
```

## Equivalence table

Function (< 2.0.0) | Equivalence (>= 2.0.0) | deprecated
--------------|----------------|-------------
`averagearea` | No equivalence, aggregating by area can be done with the `sf` package | `TRUE`
`correctionpoint`, `correctionpoints`, `correctionpoints.errors`, `correction_series` | No equivalence, better correction methods are provided by other packages (see * and * for example) | `TRUE`
`defaultCorrectionParams`, `defaultGenerationParams` | No equivalence, weather generation methods will dissapear in future versions of meteoland, as there are better alternatives (see * and *) | `TRUE`
`defaultInterpolationParams` | No change | `FALSE`
`download_*` functions | No equivalence, meteo download functions are now provided by `meteospain` package | `TRUE`
`extractdates`, `extractgridindex`, `extractgridpoints`, `extractNetCDF`, `extractvars` | No equivalence, not needed as the meteo objects are now sf objects | `TRUE`
`humidity_*` conversion tools | No change | `FALSE`
`interpolation.calibration` | `interpolator_calibration` | `TRUE`
`interpolation.calibration.fmax` | `interpolator_calibration` | `TRUE`
`interpolation.coverage` | No equivalence | `TRUE`
`interpolation.cv` | `interpolation_cross_validation` | `TRUE`
`interpolationgrid`, `interpolationpixels`, `interpolationpoints` | `interpolate_data` | `TRUE`
`mergegrid`, `mergepoints` | No equivalence, meteo objects are now sf objects and can be merged, joined or filtered as any data.frame | `TRUE`
`meteocomplete` | `complete_meteo` | `TRUE`
`meteoplot` | No equivalence, meteo objects are now sf objects and can be plotted as any other data.frame | `TRUE`
`Meteorology_*_Data` | No equivalence, spatial classes based on sp are now deprecated in `meteoland` | `TRUE`
`penman`, `penmanmonteith` | No changes | `FALSE`
`plot.interpolation.cv` | No equivalence | `TRUE`
`precipitation_concentration` | No equivalence, precipitation_concentration utility is deprecated and will be removed in future versions | `TRUE`
`precipitation_rainfallErosivity` | `precipitation_rainfall_erosivity` | `TRUE`
`radiation_*` utility functions | No change | `FALSE`
`readmeteorology*` | No equivalence, spatial classes based on sp are now deprecated in `meteoland` | `TRUE`
`readNetCDF*` | No equivalence, NetCDF files can be managed with more recent and up to date R packages (ncmeta, stars...) | `TRUE`
`readWindNinjaWindFields` | No equivalence | `TRUE`
`reshapemeteospain` | `meteospain2meteoland` | `TRUE`
`reshapeworldmet` | `worldmet2meteoland` | `TRUE`
`reshapeweathercan` | No equivalence, weathercan package was removed from CRAN | `TRUE`
`Spatial**Meteorology`, `Spatial**Topography` | No equivalence, spatial classes based on sp are now deprecated in `meteoland` | `TRUE`
`summary*` | `summarise_interpolate_data`, `summarise_interpolator` | `TRUE`
`utils_*` | No change | `FALSE`
`weathergeneration` | No equivalence, weather generation methods are deprecated | `TRUE`
`writemeteorology*` | No equivalence, spatial classes based on sp are now deprecated in `meteoland` | `TRUE`


