---
title: "Reshaping meteo data for meteoland"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reshaping meteo data for meteoland}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(meteoland)
library(stars)
library(dplyr)
```

```{r data_preparation, echo=FALSE}
unformatted_meteo <- tibble(
  date = rep(seq(from = as.Date("2022-12-01"), to = as.Date("2022-12-05"), by = 1), 3),
  station = rep(letters[1:3], each = 5),
  latitude = rep(c(41.35, 40.11, 42.00), each = 5),
  longitude = rep(c(-0.33, 0.12, 1.12), each = 5),
  min_temp = rnorm(15, 10, 6),
  max_temp = min_temp + rnorm(15, 10, 3),
  rh = rnorm(15, 50, 25)
) |>
  dplyr::mutate(
    rh = dplyr::if_else(rh > 100, 100, rh),
    rh = dplyr::if_else(rh < 20, 20, rh)
  )
```


## Meteo data format

For interpolating the meteorological variables in our locations, we need a
reference meteo data. In `meteoland` this is a `sf` object with reference
locations (usually meteorological stations) and daily values of the meteo
variables needed to perform the interpolation, *i.e.*:

```{r meteoland_meteo_example}
meteoland_meteo_example
```

`meteoland` expect names to be as in the example:

```{r meteo_names}
names(meteoland_meteo_example)
```

The only mandatory variables are `MinTemperature` and `MaxTemperature`. Other variables
(`Precipitation`, `WindSpeed`...), when present, allow for a more complete
interpolation.

## Converting meteo data to `meteoland` format

Meteo data can come in many formats and with different names for the same
variables. As we saw above, we need to convert it to a `meteoland` compliant
format (`sf` object with correct names).

### Converting from data frames

If we have our meteo data in a data.frame (*i.e.* we obtained the meteo from
our own meteo stations, or download it from other source in this format) we can
just simply transform to the desired format:

```{r unformatted_meteo}
unformatted_meteo
```

```{r unformatted_transformation}
ready_meteo <- unformatted_meteo |>
  # convert names to correct ones
  dplyr::mutate(
    MinTemperature = min_temp,
    MaxTemperature = max_temp,
    MeanRelativeHumidity = rh
  ) |>
  # transform to sf (WGS84)
  sf::st_as_sf(
    coords = c("longitude", "latitude"),
    crs = sf::st_crs(4326)
  )

ready_meteo
```

And *voilà*, we have our meteo data in the correct format

### Meteo data from other R packages

`meteoland` offers transformation functions for meteo data downloaded from the
[`meteospain`](https://emf-creaf.github.io/meteospain/) and
[`worldmet`](https://davidcarslaw.github.io/worldmet/) R packages.

#### `meteospain` data

For data coming from `meteospain` package we have the `meteospain2meteoland`
function that transforming the data for us:

```{r meteospain}
library(meteospain)
get_meteo_from(
  "meteogalicia",
  meteogalicia_options('daily', as.Date("2022-12-01"), as.Date("2022-12-05"))
) |>
  meteospain2meteoland()
```

#### `worldmet` data

For data coming from `worldmet` package we have the `worldmet2meteoland`
function that do the reshaping for us:

```{r worldmet}
library(worldmet)
worldmet::importNOAA("081120-99999", year = 2022) |>
  worldmet2meteoland()
```

### meteo data from raster sources

As we have seen, we need the meteo reference data in a `sf` object (points).
The easiest way of achieving this with meteo data in raster format is
transforming the cells to points and get the values:

TODO!!!
