---
title: "User's guide"
author: "Miquel De Cáceres, Victor Granda"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{User's guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

## Purpose

Reliable meteorological data are a basic requirement for hydrological and ecological studies at the landscape scale. Given the large spatial variation of meteorology over complex terrains, meteorological records from a single weather station are often not representative of entire landscapes. Studies made on multiple sites over a landscape require different meteorological series for each site; and other studies may require meteorological data series for all grid cells of a landscape, in a continuous way. In these cases, spatial correlation between the meteorology series of different sites or cells must be taken into account. For example, the sequence of days with rain of contiguous cells will normally be the same or very similar, even if precipitation amounts may differ. Finally, studies addressing the impacts of climate change on forests and landscapes require downscaling coarse-scale predictions  of global or regional climate models to the landscape scale. When downscaling predictions for several locations in a landscape, spatial correlation of predictions is also important.

With the aim to assist research of climatic impacts on forests, the R package `meteoland` provides utilities to estimate daily weather variables at any position over complex terrains:

1. Spatial interpolation of daily weather records from meteorological stations. }
2. Statistical correction of meteorological data series (e.g. from climate models).}

Using `meteoland` package is easy, but some ideas and concepts must be addresed
to make the experience easy for new users. This vignette is intended as a
working example to explain all the steps needed to get comfortable with
`meteoland` workflow. We assume that the reader is more or less familiar with S4 classes of package **sp**.

## Installing and loading the package

First of all, before starting to work with the package, we must install and load
the library. `meteoland` stable version can be found at CRAN
(https://cran.r-project.org/web/packages/meteoland/index.html), and it can be
installed and loaded as any other R package:

```{r, eval = FALSE}
install.packages("meteoland")
library(meteoland)
```

Alternatively, you can install the development version located at GitHub using the `devtools` package:
```{r, eval = FALSE}
library(devtools)
install_github("miquelcaceres/meteoland", build_vignettes = TRUE)
library(meteoland)
```


# Interpolation of daily weather
## Preparing weather data for interpolation


```{r, echo = FALSE, include= FALSE}
library(meteoland)
data("exampleinterpolationdata")

st_names = row.names(exampleinterpolationdata@MinTemperature)
st_coords = exampleinterpolationdata@coords
st_data = data.frame(X_UTM = st_coords[,1], Y_UTM = st_coords[,2],
                     elevation = exampleinterpolationdata@elevation,
                     row.names = st_names)
tmin = exampleinterpolationdata@MinTemperature
tmax = exampleinterpolationdata@MaxTemperature                     
prec = exampleinterpolationdata@Precipitation                     
```

Before starting using the package, you need to have access to the elevation (in m) and daily weather data series corresponding a set of locations (normally weather stations). Elevation is needed because interpolation routines perform corrections for differences in elevation between the reference locations and the target point. The initial format of your data will be different depending on the format used by your data provider (the package has also tools to access weather data). For our example, we will assume you have data from a set of 38 stations in your study area. On one hand, you should have a `data.frame` with the coordinates and elevation of each location:
```{r}
str(st_data)
head(st_data)
```
On the other, you should have at least three matrices of meteorological data (one for minimum temperature, one for maximum temperature and the last one for precipitation) with stations in rows and dates in columns:
```{r}
dim(tmax)
dim(tmin)
dim(prec)
tmax[1:6,1:6]
```
Units should be in degrees Celsius for temperature and mm for precipitation.

## Building an interpolation data object

Meteoland stores weather series for reference locations and interpolation parameters in a single object of class `MeteorologyInterpolationData`. There are several ways of building such objects, but we will first illustrate how to do it from the data we just presented.

First we need to create an object of class `SpatialPoints` (see package **sp**) with the spatial coordinates of our stations and the coordinate system (here UTM 31N):
```{r}
sp = SpatialPoints(st_data[,c("X_UTM", "Y_UTM")],
                   proj4string = CRS("+proj=utm +zone=31 +ellps=WGS84 +datum=WGS84 +units=m +towgs84=0,0,0 "))
head(sp)
```
We can now build an object `MeteorologyInterpolationData` using:
```{r}
interpolator <- MeteorologyInterpolationData(sp, elevation = st_data$elevation,
                                             MinTemperature = tmin,
                                             MaxTemperature = tmax,
                                             Precipitation = prec)
class(interpolator)
```

The resulting object is now ready to be used to perform interpolation on a set of target locations (see next section). We can inspect the amount of data in our interpolation object using:
```{r}
spatial_coverage <- interpolation.coverage(interpolator, type = 'spatial')
head(spatial_coverage)
temporal_coverage <- interpolation.coverage(interpolator, type = 'temporal')
head(temporal_coverage)
```
In the first call, function `interpolation.coverate` returns the number of non-missing observations for each station and variable. The second call produces a similar output, but by dates instead of stations.

Interpolation parameters are also stored in the same object (do not worry if you do not understand their meaning now):
```{r}
names(interpolator@params)
```

An important parameter at this point is the one specifying an initial radius for the truncated spatial Gaussian kernel, which in this case is:
```{r}
interpolator@params$initial_Rp
```

Its value must be set in relation to the units of the spatial coordinates of weather data. In this case, our data was in meters, so this radius indicates 140 km. In general, the initial radius should be enough to include a reasonable number of stations (~20-40), but the kernel radius is adjusted for each interpolation target point.


## Calibration and cross-validation of the interpolation data
Once we already have weather stations data in shape, we can start calibrating the model in order to obtain the optimal parameters for the meteorological variables we want to interpolate. Parametere calibration has to be done for each variable separately. For example for minimum temperature:
```{r}
tmin_cal <- interpolation.calibration(interpolator, variable = "Tmin",
                                      N_seq = 20,
                                      alpha_seq = seq(5, 10, by = 1),
                                      verbose = TRUE)
```

This function returns an interpolation.calibration class object which contains several items:

* Numeric matrix with the mean absolute error (MAE) values for each combination of parameters $N$ and
$\alpha$.
* Miminum value found for MAE.
* Value for the $N$ parameter corresponding to the minumun MAE.
* Value for the $\alpha$ parameter corresponding to the minimum MAE.
* Matrix with the observed values.
* Matrix with the predicted values for the optimum parameter combination.

The result of the calibration needs to be manually stored in the interpolation params:
```{r}
interpolator@params$N_MinTemperature = tmin_cal$N
interpolator@params$alpha_MinTemperature = tmin_cal$alpha
```

In general, we recommend conducting calibration exercises at least once for each variable and each data set used as reference for interpolation, and more than once periods differ in the number of stations available.

Before using the object for interpolations, we also need to assess its performance. This is done by leave-one-out cross-validation in function `interpolation.cv`:
```{r,echo=FALSE}
cv <- interpolation.cv(interpolator, verbose=T)
```

And the results are inspected using:
```{r}
summary(cv)
```

## Interpolation on a grid
```{r, echo = FALSE, include= FALSE}
data("examplegridtopography")

elev = as(examplegridtopography[1:20,1:20], "SpatialGridDataFrame")["elevation"]
```
The target for weather interpolation in `meteoland` can be a set of points, pixels or a whole grid. Again, the initial format of data can be very different. Here we assume you have a small grid of 400 (20x20) cells of 1ha in size, with elevation data in form of class `SpatialGridDataFrame` (see method `read.asciigrid` in package *sp*):
```{r}
summary(elev)
```

Note that the coordinate reference system needs to be the same as that of `interpolator`, which in this case it is. Before performing the interpolation over this grid, we need to reshape this data in a class called `SpatialGridTopography`:
```{r}
sgt = SpatialGridTopography(as(elev, "SpatialGrid"), elevation = elev$elevation,
                            proj4string = elev@proj4string)
sgt
```

As you can see in the result, `meteoland` has calculated for us slope and aspect (both in degrees) from elevation data ^[Objects of class `SpatialGridTopography` can be initialized with user input values for slope and aspect too, but `meteoland` has its own routines when this are missing]. Slope and aspect are important for radiation calculations, which also requires relative humidity data. We can display elevation over the grid using:
```{r}
spplot(sgt, "elevation")
```


Before we call the interpolation routine, we need to define the dates (i.e. days) for which we want weather to be interpolated, for example:
```{r}
dates = as.Date(c("2001-02-03", "2001-06-03"))
```

Of course, we need to be sure that the `interpolator` object has data corresponding to this dates. We can check if there is any missing date using:
```{r}
sum(!(dates %in% interpolator@dates))
```

The name of interpolation functions depend on the target spatial structure. For grids we need to use function `interpolationgrid`:
```{r}
ml <- interpolationgrid(interpolator, sgt, dates)
```

This function works processing each date at a time. Since calculations can take some time, the console output shows the progress. The output of the function is an object of class `SpatialGridMeteorology`:
```{r}
ml
```


We can display the result in a map using function:
```{r, fig.show="hold"}
spplot(ml, 2, "MinTemperature")
spplot(ml, 2, "MaxTemperature")
```

## Interpolation on a set of points
```{r, echo = FALSE, include= FALSE}

spt = as(examplegridtopography, "SpatialPointsTopography")[c(36,228, 530, 2304)]
cc = coordinates(spt)
points_df = data.frame(X_UTM = cc[,1], Y_UTM = cc[,2], elevation = spt@data$elevation)
```
If you want to interpolate on a set of target locations, the starting point will normally be a `data.frame`. In our example this points come from the grid, but we have reshaped them so the starting format is familiar:
```{r}
points_df
```

Analogously with the grid, we need to transform this data into an object `SpatialPointsTopography`:
```{r}
spt = SpatialPointsTopography(as.matrix(points_df[,c("X_UTM", "Y_UTM")]),
                              elevation = points_df$elevation,
                              proj4string = CRS("+proj=utm +zone=31 +ellps=WGS84 +datum=WGS84 +units=m +towgs84=0,0,0"))
spt
```

In this case we only have elevation (in m), but slope and aspect should also be included if possible. Let us assume you want to interpolate on this points for the whole time series available in object `interpolator`. Since we are dealing with points, the function to interpolate is called `interpolationpoints`:
```{r}
mp = interpolationpoints(interpolator, spt)
```
This function works processing one point at a time. The output of the function is an object of class `SpatialPointsMeteorology`:
```{r}
mp
```

And the precipitation series for point #1 can be plotted using:
```{r, fig =TRUE, fig.width=6}
meteoplot(mp, 1, "Precipitation", ylab="Precipitation (mm)", xlab="")
```

