---
title: "Interpolation over Senegal"
author: "Miquel De Cáceres"
date: "6 de gener de 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(meteoland)
```

In this document we show how to obtain and process the topographic and weather data needed by package **meteoland** to conduct weather interpolation over a given country. We'll use Senegal as case study.

## Accessing weather station data

We start by loading package **worldmet**, which provides access to over 30.000 weather stations across the globe:
```{r}
library(worldmet)
```
In order to download weather data, we must first obtain the codes of the weather stations available for our study area. This can be done using function `getMeta()` from worldmet, which we use to know available weather stations for Senegal and neighbouring countries:
```{r}
#Country codes for Senegal / Gambia/ Guinea/ Mali/ Mauritania /Guinea bissau
countries <- c("SG", "GA", "GV", "ML", "MR", "PU")
sg_st = getMeta(country=countries, plot = FALSE) 
sg_st
```
Let's assume that we want to interpolate weather for year 2018. We can download hourly weather station data for the selected stations and year 2018 using function `importNOAA()`:
```{r}
hour2018<-importNOAA(code = sg_st$code, year=2018, precip= TRUE)
```
Some stations/years were missing. Using function `reshapeworldmet()` from meteoland, we can reshape downloaded hourly data into daily data in form of an object `SpatialPointsMeteorology`:
```{r}
spt_sen<-reshapeworldmet(hour2018, output = "SpatialPointsMeteorology", verbose=FALSE)
spt_sen
```


## Accessing elevation data

We can use function `getData()` from package **raster** to download elevation data for Senegal:
```{r, echo=FALSE}
library(raster)
elev_raster <- getData('alt', country="SEN")
elev_raster
```
The previous function returns a `Raster` grid object in geographic coordinates, which we can plot along with the station locations:
```{r}
plot(elev_raster)
plot(spt_sen, add=TRUE)
```

Reproject the raster to 29N UTM to have meters as coordinate units
```{r}
elev_utm = projectRaster(elev_raster, crs = CRS("+init=epsg:32628"))
elev_utm
```
Reshape into a `SpatialGridDataFrame`:
```{r}
elev_sgd = as(elev_utm, "SpatialGridDataFrame")
```

Build object of `SpatialGridTopography`
```{r}
sgt <- SpatialGridTopography(elev_sgd@grid, elevation = elev_sgd$SEN_msk_alt, 
                             proj4string = proj4string(elev_sgd))
spxt <- as(sgt, "SpatialPixelsTopography")
```

Check slope and aspect calculated by meteoland
```{r, fig=TRUE}
spplot(spxt, "slope")
```

```{r, fig=TRUE}
spplot(spxt, "aspect")
```

## Performing interpolation

```{r}
spt_mid<-reshapeworldmet(hour2018, output = "MeteorologyInterpolationData", 
                         proj4string = CRS("+init=epsg:32628"),
                         verbose=FALSE)
```

```{r}
mid <- interpolationpixels(spt_mid, spxt, dates=as.Date(c("2018-03-15", "2018-07-30")))
```

```{r, fig=TRUE}
spplot(mid,1,"MeanTemperature")
```

```{r, fig=TRUE}
spplot(mid,1,"MeanRelativeHumidity")
```
