---
title: "Interpolation over Senegal"
author: "Miquel De Cáceres"
date: "6 de gener de 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(meteoland)
```

## Accessing weather station data


```{r}
library(worldmet)
```

```{r}
sg_st = getMeta(country="SG", plot = FALSE) #Senegal
ga_st = getMeta(country="GA", plot = FALSE) #Gambia
gv_st = getMeta(country="GV", plot = FALSE) #Guinea
ml_st = getMeta(country="ML", plot = FALSE) #Mali
mr_st = getMeta(country="MR", plot = FALSE) #Mauritania
pu_st = getMeta(country="PU", plot = FALSE) #Guinea-Bissau
codes <- c(sg_st$code, ga_st$code, gv_st$code, ml_st$code, mr_st$code, pu_st$code)
```

Download hourly weather station data for year 2018
```{r}
hour2018<-importNOAA(code = codes, year=2018, precip= TRUE)
```
Reshape downloaded hourly data into daily data in form of an object `SpatialPointsMeteorology`:
```{r}
spt_sen<-reshapeworldmet(hour2018, output = "SpatialPointsMeteorology", verbose=FALSE)
spt_mid<-reshapeworldmet(hour2018, output = "MeteorologyInterpolationData", verbose=FALSE)
```


## Accessing elevation data

Download elevation data for senegal
```{r, echo=FALSE}
library(raster)
elev_raster <- getData('alt', country="SEN")
```


```{r}
plot(elev_raster)
plot(spt_sen, add=TRUE)
```

Reproject raster into UTM to have meters as coordinate units
```{r}
elev_utm = projectRaster(elev_raster, crs = CRS("+init=epsg:32628"))
```
Reshape into a `SpatialGridDataFrame`:
```{r}
elev_sgd = as(elev_utm, "SpatialGridDataFrame")
```

Build object of `SpatialGridTopography`
```{r}
sgt <- SpatialGridTopography(elev_sgd@grid, elevation = elev_sgd$SEN_msk_alt, 
                             proj4string = proj4string(elev_sgd))
```

Check slope and aspect calculated by meteoland
```{r, fig=TRUE}
spplot(sgt, "slope")
```

```{r, fig=TRUE}
spplot(sgt, "aspect")
```

## Performing interpolation

```{r}
mid <- interpolationgrid(spt_mid, grid = sgt, dates=as.Date(c("2018-03-15", "2018-07-30")))
```

```{r, fig=TRUE}
spplot(mid,1,"MeanTemperature")
```

```{r, fig=TRUE}
spplot(mid,2,"MeanTemperature")
```
