---
title: "Daily weather interpolation over Senegal"
author: "Miquel De Cáceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: false
params:
  complete_rebuild: FALSE
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we show how to obtain and process the topographic and weather data needed by package **meteoland** to conduct weather interpolation over a given country, using package **worldmet** as source for reference weather station data. We'll use Senegal as case study.

We begin by loading necessary libraries, all available from CRAN:
```{r, include=TRUE, message= FALSE, echo = TRUE}
library(sf)
library(dplyr)
library(terra)
library(stars)
library(geodata)
library(tidyterra)
library(meteoland)
library(ggplot2)
library(cowplot)
```

We will also use package **afrilearndata** from GitHub, to display African country limits:
```{r, message= FALSE, echo = TRUE}
# remotes::install_github("afrimapr/afrilearndata")
library(afrilearndata)
```

## Accessing weather station data

We start by loading package **worldmet**, which provides access to over 30.000 weather stations across the globe:
```{r, echo = TRUE}
library(worldmet)
```

In order to download weather data, we must first obtain the codes of the weather stations available for our study area. This can be done using function `getMeta()` from worldmet, which we use to know available weather stations for Senegal and neighbouring countries:

```{r, eval = params$complete_rebuild}
# Country codes:
#   Senegal - SG
#   Gambia - GA 
#   Guinea - GV
#   Mali - ML 
#   Mauritania - MR
#   Guinea bissau - PU
sg_st <- dplyr::bind_rows(worldmet::getMeta(country="SG", plot = FALSE),
                          worldmet::getMeta(country="GA", plot = FALSE),
                          worldmet::getMeta(country="GV", plot = FALSE),
                          worldmet::getMeta(country="ML", plot = FALSE),
                          worldmet::getMeta(country="MR", plot = FALSE),
                          worldmet::getMeta(country="PU", plot = FALSE))
```

```{r, eval = TRUE, echo=FALSE}
if(params$complete_rebuild) {
  saveRDS(sg_st, file="sg_st.rds")
} else {
  sg_st <- readRDS("sg_st.rds")
}
```

The station metadata looks like this:
```{r}
sg_st
```

Let's assume that we want to interpolate weather for year 2020. We can download hourly weather station data for the selected stations and year 2020 using function `importNOAA()` and programming a loop over stations:

```{r, eval = params$complete_rebuild, echo = TRUE, include = TRUE}
hourData <- NULL
for(i in 1:length(sg_st$code)) {
  cat(paste0(sg_st$code[i]," "))
  tryCatch({
    hi <- worldmet::importNOAA(code = sg_st$code[i], year=2020, quiet=TRUE)
    if(is.null(hourData))  {hourData <- hi}
    else {hourData <- dplyr::bind_rows(hourData,hi)}
    cat(" [success].")
  }, error = function(e) {cat("[failed].")})
}
```

```{r, eval = TRUE, echo=FALSE}
if(params$complete_rebuild) {
  saveRDS(hourData, file="hourData.rds")
} else {
  hourData <- readRDS("hourData.rds")  
}
```

## Processing weather station data for interpolation

Some stations had missing data for year 2020 and were excluded from the output. Using function `worldmet2meteoland()` from **meteoland**, we can reshape the hourly data of the remaining stations into daily data in form of an `sf` object:

```{r, eval = params$complete_rebuild, echo = FALSE}
sf_sen <- worldmet2meteoland(hourData, complete = TRUE)
sf_sen
```

```{r, eval = TRUE, echo=FALSE}
if(params$complete_rebuild) {
  saveRDS(sf_sen, file="sf_sen.rds")
} else {
  sf_sen <- readRDS("sf_sen.rds")  
}
```

An interpolator object needs to be created for interpolation, containing not only station weather and topographic data but also interpolation parameters. This can be obtained using function `create_meteo_interpolator()`:

```{r}
interpolator_sen <- meteoland::create_meteo_interpolator(sf_sen, verbose = FALSE)
```

The warning message indicates that default interpolation parameters are set and in a serious application we should calibrate the parameters before using the reference data for interpolation. 

The interpolator object is of class **stars**:
```{r}
interpolator_sen
```

If needed, we can display the location of the weather stations on a map with country limits. This is useful to know the station density in the study area. In our case we will use the following code:

```{r, fig.width=4, fig.height=6}
data("africountries")
senegal <- africountries[africountries$name=="Senegal",]
ggplot()+
  geom_sf(data = sf::st_geometry(africountries))+
  geom_sf(data = sf::st_geometry(senegal), fill= "red")+
  geom_sf(data = sf_sen[1], size = 0.5)+
  xlim(c(-20,0))+ ylim(c(0,30))+
  theme_bw()
```


## Accessing and arranging topographic data

To perform weather interpolation, we need the topography (elevation, slope, aspect) of the target area. We can use function `elevation_30s()` from package **geodata** to download elevation data for Senegal:

```{r, echo=FALSE}
elev_raster <- geodata::elevation_30s(country="SEN", path = tempdir())
names(elev_raster) <- "elevation"
elev_raster
```

The previous function return `SpatRaster` object (package **terra**). In order to estimate slope and aspect we can use function `terrain()` from the **terra** package:
```{r}
slope_raster <- terra::terrain(elev_raster, v = "slope", unit="degrees")
aspect_raster <- terra::terrain(elev_raster, v = "aspect", unit="degrees")
```

We can plot the three rasters using with `ggplot()` using `geom_spatraster()` from package **tidyterra**:

```{r, fig.width=4, fig.height=8}
p1 <- ggplot()+
  tidyterra::geom_spatraster(data = elev_raster)
p2 <- ggplot()+
  tidyterra::geom_spatraster(data = slope_raster)
p3 <- ggplot()+
  tidyterra::geom_spatraster(data = aspect_raster)
cowplot::plot_grid(p1, p2, p3, nrow=3)
```



We now assemble the three rasters into a single object:

```{r}
topo_spatraster <- c(elev_raster, slope_raster, aspect_raster)
```

The resolution of the raster is quite high, which would make interpolation quite slow. For this reason,
we use function `aggregate()` from package **terra** to lower the resolution and speed-up calculations, but one may skip this step.
```{r}
fact <- 10
topo_spatraster_agg <- terra::aggregate(topo_spatraster, fact = fact)
```

Finally, we reshape the raster into an object of package **stars**:

```{r}
topo_stars_agg <- stars::st_as_stars(topo_spatraster_agg, as_attributes=TRUE)
```

## Performing interpolation

We are now ready to perform the interpolation of weather over the study area. We can simply call function `interpolate_data()` from **meteoland**. Here we restrict the interpolation to two dates in 2020 to speed up calculations:
```{r}
raster_interpolated <- topo_stars_agg |>
  interpolate_data(interpolator_sen, dates = c(as.Date("2020-01-01"), as.Date("2020-07-01")), 
                   verbose = FALSE)
```
Warnings are raised because some raster locations (those over the ocean) are outside the convex hull of the stations in the interpolator object. 

The result of the interpolation is a raster object of class **stars**:
```{r}
raster_interpolated
```

We can display the maps for specific dates and variables using:
```{r, fig.width=7, fig.height=8}
p1 <- ggplot()+
      geom_stars(data = raster_interpolated["MeanTemperature",,,1])+
      scale_fill_viridis_c()+
      labs(title = "2020-01-01")
p2 <- ggplot()+
      geom_stars(data = raster_interpolated["MeanTemperature",,,2])+
      scale_fill_viridis_c()+
      labs(title = "2020-07-01")
cowplot::plot_grid(p1, p2, nrow = 2)
```

